---
title: "Japan suicide cases reported by National Police Agency"
author: "Mitsuo Shiota"
date: "2020-12-19"
output: 
  github_document:
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  fig.width = 7,
  fig.asp = 0.618,
  out.width = "90%"
  )

library(tidyverse)
library(lubridate)
library(readxl)
library(tsibble)
library(tqr)
library(scales)
```

Updated: `r Sys.Date()`

## Motivation

I was watching News Center 9, an NHK nightly news program, on December 16, 2020. It told suicide cases have been rapidly surging since July 2020 in Japan by showing a chart, which I will replicate later. I felt the apparent rapid surge might be an exaggeration. 

## Get data from National Police Agency

I make a csv file manually from the data in [National Police Agency site](https://www.npa.go.jp/publications/statistics/safetylife/jisatsu.html). I can get monthly suicide cases, not separated by gender but total, since January 2008. 

```{r input, message=FALSE}
npa <- read_csv("data/npa.csv")
npa_data <- npa %>% 
  pivot_longer(!year, names_to = "month", values_to = "cases") %>% 
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    time = make_date(year, month, 01L)
  ) %>% 
  arrange(time)

head(npa_data)
```

## Replicate the chart

I watch the program again in [NHK plus site](https://plus.nhk.jp/), which requires registration, and replicate the chart below, though the original chart did not include November and December 2020. Look at the 2020 line, and find it exceeds the Mean (2017-2019) line since July.

Yes, the 2020 line shows a rapid surge up to October. But, why was it significantly below the Mean (2017-2019) line from April to June in the first place? I remember National Declaration of State of Emergency to fight Covid-19 was effective from April 15 to May 24. As "month" indicates the timing of discovery, not of commitment, in the National Police Agency data, I suspect some cases were not discovered in the emergency period, and were discovered later. This may explain some part of excess from July to October.

```{r replicate, warning=FALSE}
df_last_3_yr_avg <- npa_data %>% 
  filter(time >= "2017-01-01", time <= "2019-12-01") %>% 
  select(!time) %>% 
  pivot_wider(names_from = month, values_from = cases) %>% 
  summarize(across(!year, mean)) %>% 
  pivot_longer(everything(), names_to = "month", values_to = "last_3_yr_avg") %>% 
  mutate(month = as.integer(month))

df_10_12 <- npa_data %>% 
  filter(time >= "2020-10-01", time <= "2020-12-01") %>% 
  select(c(month, cases))

npa_data %>% 
  filter(time >= "2020-01-01", time <= "2020-10-01") %>% 
  select(c(month, cases)) %>% 
  right_join(df_last_3_yr_avg, by = "month") %>% 
  rename(
    `2020` = cases,
    `Mean (2017-2019)` = last_3_yr_avg
  ) %>% 
  pivot_longer(!month, names_to = "year", values_to = "cases") %>% 
  ggplot() +
  geom_line(aes(month, cases, color = year)) +
  geom_line(data = df_10_12, aes(month, cases), linetype = "dashed", color = "red") +
  scale_color_manual(values = c("red", "blue")) +
  scale_x_continuous(breaks = 1:12) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Suicide cases",
    y = NULL, color = NULL
    )
```

## Suicide cases are bouncing back, but not as fast as the program suggests

As I read the book "Towards Evidence-Based Suicide Prevention: Perspectives from Economics and Political Science", co-authored by the original chart creator in the program, I know that the unemployed are more likely to commit suicide than the employed, and that unemployment rates are highly correlated with suicide cases.

As Japan has been officially in recession since October 2018, the unemployment rates hit the bottom in late 2019, and began to climb in 2020.

```{r recession}
recessions_df = tribble(
  ~Peak, ~Trough,
  #----------|----------
  # "1951-06-01", "1951-10-01",
  # "1954-01-01", "1954-11-01",
  # "1957-06-01", "1958-06-01",
  # "1961-12-01", "1962-10-01",
  # "1964-10-01", "1965-10-01",
  # "1970-07-01", "1971-12-01",
  # "1973-11-01", "1975-03-01",
  # "1977-01-01", "1977-10-01",
  # "1980-02-01", "1983-02-01",
  # "1985-06-01", "1986-11-01",
  # "1991-02-01", "1993-10-01",
  # "1997-05-01", "1999-01-01",
  # "2000-11-01", "2002-01-01",
  "2008-02-01", "2009-03-01",
  "2012-03-01", "2012-11-01",
  "2018-10-01", "2021-04-01"
)

recessions_df <- recessions_df %>% 
  mutate(
    Peak = as.Date(Peak),
    Trough = as.Date(Trough)
  )

```


```{r unemployment_chart, message=FALSE, results=FALSE}
tf <- tempfile(fileext = ".xlsx") # lt01-a10.xlsx

url <- "https://www.e-stat.go.jp/stat-search/file-download?statInfId=000031831358&fileKind=0"

httr::GET(url, httr::write_disk(tf))

unemployment_rate <- read_excel(tf, col_names = FALSE, skip = 9)
unemployment_rate <- unemployment_rate[, c(1:2, 20)]
names(unemployment_rate) <- c("year", "month", "percent")

unemployment_rate$time <- (as.Date("1953-01-01") + months(0:(nrow(unemployment_rate) - 1)))

unemployment_rate <- unemployment_rate %>% 
  drop_na(percent)

unemployment_rate %>% 
  select(c(time, percent)) %>% 
  filter(time >= "2008-01-01") %>% 
  ggplot(aes(time, percent)) +
  geom_line() +
  geom_rect(data = recessions_df, inherit.aes = FALSE,
            aes(xmin = Peak, xmax = Trough, ymin = -Inf, ymax = +Inf),
            fill='darkgray', alpha=0.5) +
  labs(
    title = "Unemployment rates (seasonally-adjusted)",
    x = NULL
  )

```


I draw a chart of both original and seasonally-adjusted cases since 2008. Suicide cases indeed hit the bottom in the beginning of 2020, and are bouncing back, even though I consider the fluctuations in 2020 should be smoothed out.

```{r simple_chart, fig.asp=0.7}
npa_ts <- npa_data %>% 
  drop_na(cases) %>% 
  select(c(time, cases)) %>% 
  mutate(
    key = "original",
    time = yearmonth(time)
    ) %>% 
  as_tsibble(key = key, index = time)

npa_sa <- npa_ts %>% 
  tq_sa() %>% 
  mutate(key = "seasonally adjusted")

npa_ts %>% 
  bind_rows(npa_sa) %>% 
  ggplot(aes(time, cases, color = key)) +
  geom_line() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Suicide cases",
    x = NULL, y = NULL, color = NULL
    ) +
  theme(legend.position = "bottom")
  
```

EOL